Outbreak response models of vaccine-preventable diseases in humans (1970-2019) - A systematic review (Supplementary Files)

### 2023-06-21 analyses/scripts/orv_sys_rev_supplementary.Rmd
Renamed the data object

# Filter the FMD data
fmd_data <- compact_data_with_citation_keys %>% 
  filter(disease == 'fmd') 
```

@@

```{r echo=FALSE}
#Cross tally the model structure with parametrization and validation categories
#FIXME: Check that studies are either not being under counted or double-counted. What do I want to count here? Is it the unique studies per method or combined method because most studies used a combination of parametrization methods especially
parametrization_and_validation_by_model_structure <- compact_data_no_fmd %>%
  mutate(parametrization = str_replace(parametrization, ' ', ' + '),
         validation = str_replace(validation, ' ', ' + ')
         ) %>% 
  count(model_structure, parametrization, validation) %>% 
  pivot_wider(names_from = model_structure, values_from = n)

@@

```{r echo=FALSE}
#Keep only the unique entries
parametrization_and_validation_count <- compact_data_no_fmd %>% 
  mutate(parametrization = str_replace(parametrization, ' ', ' + '), 
         validation = str_replace(validation, ' ', ' + ')
         ) %>% 
  count(parametrization, 
        validation, 
        name = 'num_of_studies'
        ) %>% 
  group_by(parametrization) %>% 
  arrange(desc(num_of_studies), .by_group = T)

@@

```{r echo=FALSE}
#Keep only the unique entries
code_and_data_availability_count <- compact_data_no_fmd %>% 
  count(data_used, data_available) %>% 
  pivot_wider(names_from = data_available, values_from = n) %>% 
  replace_na(list(not_applicable = 0, 
                  yes = 0, 
                  no = 0
                  )
             )

@@

```{r echo=FALSE}
#Keep only the unique entries
code_availability_count <- compact_data_no_fmd %>% 
  count(simulation_code_available)

@@

```{r echo=FALSE}
#count the number of unique diseases studied
disease_studied_by_outbreak_type <- compact_data_no_fmd %>% 
  separate_rows(disease,  sep = ' ') %>% 
  mutate(disease = as_factor(disease), 
         outbreak_type = as_factor(str_replace(outbreak_type, ' ', '+'))
         ) %>% 
  group_by(disease) %>% 
  distinct(title, .keep_all = TRUE) %>% 
  count(disease, outbreak_type, sort = TRUE) %>%
  pivot_wider(names_from = outbreak_type, values_from = n) %>% 
  replace_na(list(hypothetical_outbreak = 0, real_outbreak = 0)) %>% 
  mutate(total = hypothetical_outbreak + real_outbreak) %>% 
  arrange(desc(total))

@@

```{r echo=FALSE}
#count the number of unique diseases studied
disease_studied_by_objective_type <- compact_data_no_fmd %>% 
  separate_rows(disease,  sep = ' ') %>% 
  mutate(disease = as_factor(disease), 
         objectives = as_factor(objectives)
         ) %>% 
  group_by(disease) %>% 
  distinct(title, .keep_all = TRUE) %>% 
  count(disease, objectives) %>%
  pivot_wider(names_from = objectives, values_from = n) %>% 
  replace_na(list(future = 0, past = 0, both = 0)) %>% 
  mutate(total = future + past + both) %>% 
  arrange(desc(total))

@@

## FMD by collaboration type
```{r echo=FALSE}
fmd_studies_by_collab_type <- compact_data %>% 
  filter(disease == 'fmd') %>% 
  mutate(disease = as_factor(disease), 
         author_affiliation_type = as_factor(str_replace(author_affiliation_type, ' ', '+'))
         ) %>% 
  mutate(collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'purely_academic', 
                                         'mixed')
                                 )
         ) %>% 
  count(collab_type, sort = TRUE)

knitr::kable(fmd_studies_by_collab_type)
```
## FMD by objective type
```{r echo=FALSE}
fmd_studies_by_objective_and_outbreak_type <- compact_data %>% 
  filter(disease == 'fmd') %>% 
  count(objectives, outbreak_type) %>% 
  pivot_wider(names_from = outbreak_type, values_from = n)

@@

## Vaccine impact categories (FMD)
```{r echo=FALSE}
vax_impact_categories_fmd <- compact_data %>%
  filter(disease == 'fmd') %>% 
  filter(vax_modelled_with_non_vax == 'yes') %>% 
  count(is_vax_effective, 
        name = 'num_of_studies', 
        sort = T
        )

@@

```{r echo=FALSE}
#' Count the number of each author affiliation type per year (including FMD)
collap_types_by_year <- compact_data_no_fmd %>% 
  mutate(year_aggreg = as.numeric(ifelse(publication_year < 2006, 
                                         2005, 
                                         publication_year
                                         )
                                  ), 
         collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'Purely academic', 
                                         'Mixed')
                                 )
         ) %>% 
  count(year_aggreg, 
        collab_type, 
        name = 'total_publications'
        ) %>% 
  pivot_wider(names_from = collab_type, values_from = total_publications)

@@

```{r echo=FALSE}
academic_collabs_versus_others <- compact_data_no_fmd %>% 
  filter(author_affiliation_type != 'academic_institutions') %>% 
  mutate(year_aggreg = as.numeric(ifelse(publication_year < 2006, 2005, publication_year
                                         )
                                  ), 
         collab_type = as_factor(if_else(str_detect(author_affiliation_type, 'academic_institutions'), 
                                         'Academic', 
                                         'Other'
                                         )
                                 )
         ) %>% 
  count(year_aggreg, 
        collab_type, 
        name = 'total_publications'
        )

@@

# Connection of authors to the studied locations 
```{r}
author_location_count_by_disease <- compact_data_no_fmd %>%
  separate_rows(disease, sep = ' ') %>% 
  mutate(author_in_country_studied = as_factor(author_in_country_studied), 
         disease = as_factor(disease)
         ) %>% 
  group_by(collab_type, disease) %>%
  distinct(title, .keep_all = TRUE) %>% 
  count(author_in_country_studied, sort = TRUE) %>% 
  complete(author_in_country_studied, fill = list(n = 0)) %>% 
  pivot_wider(names_from = author_in_country_studied, values_from = n) %>% 
  replace_na(list(yes = 0, no = 0, not_applicable = 0)) %>% 
  mutate(total = not_applicable + no + yes) %>% 
  ungroup(collab_type) %>% 
  arrange(desc(total))

>>>>

# Filter the FMD data
data_fmd_only <- compact_data_with_citation_keys %>% 
  filter(disease == 'fmd') 
```

@@

```{r echo=FALSE}
#Cross tally the model structure with parametrization and validation categories
#FIXME: Check that studies are either not being under counted or double-counted. What do I want to count here? Is it the unique studies per method or combined method because most studies used a combination of parametrization methods especially
parametrization_and_validation_by_model_structure <- data_without_fmd %>%
  mutate(parametrization = str_replace(parametrization, ' ', ' + '),
         validation = str_replace(validation, ' ', ' + ')
         ) %>% 
  count(model_structure, parametrization, validation) %>% 
  pivot_wider(names_from = model_structure, values_from = n)

@@

```{r echo=FALSE}
#Keep only the unique entries
parametrization_and_validation_count <- data_without_fmd %>% 
  mutate(parametrization = str_replace(parametrization, ' ', ' + '), 
         validation = str_replace(validation, ' ', ' + ')
         ) %>% 
  count(parametrization, 
        validation, 
        name = 'num_of_studies'
        ) %>% 
  group_by(parametrization) %>% 
  arrange(desc(num_of_studies), .by_group = T)

@@

```{r echo=FALSE}
#Keep only the unique entries
code_and_data_availability_count <- data_without_fmd %>% 
  count(data_used, data_available) %>% 
  pivot_wider(names_from = data_available, values_from = n) %>% 
  replace_na(list(not_applicable = 0, 
                  yes = 0, 
                  no = 0
                  )
             )

@@

```{r echo=FALSE}
#Keep only the unique entries
code_availability_count <- data_without_fmd %>% 
  count(simulation_code_available)

@@

```{r echo=FALSE}
#count the number of unique diseases studied
disease_studied_by_outbreak_type <- data_without_fmd %>% 
  separate_rows(disease,  sep = ' ') %>% 
  mutate(disease = as_factor(disease), 
         outbreak_type = as_factor(str_replace(outbreak_type, ' ', '+'))
         ) %>% 
  group_by(disease) %>% 
  distinct(title, .keep_all = TRUE) %>% 
  count(disease, outbreak_type, sort = TRUE) %>%
  pivot_wider(names_from = outbreak_type, values_from = n) %>% 
  replace_na(list(hypothetical_outbreak = 0, real_outbreak = 0)) %>% 
  mutate(total = hypothetical_outbreak + real_outbreak) %>% 
  arrange(desc(total))

@@

```{r echo=FALSE}
#count the number of unique diseases studied
disease_studied_by_objective_type <- data_without_fmd %>% 
  separate_rows(disease,  sep = ' ') %>% 
  mutate(disease = as_factor(disease), 
         objectives = as_factor(objectives)
         ) %>% 
  group_by(disease) %>% 
  distinct(title, .keep_all = TRUE) %>% 
  count(disease, objectives) %>%
  pivot_wider(names_from = objectives, values_from = n) %>% 
  replace_na(list(future = 0, past = 0, both = 0)) %>% 
  mutate(total = future + past + both) %>% 
  arrange(desc(total))

@@

## FMD by collaboration type
```{r echo=FALSE}
fmd_studies_by_collab_type <- data_fmd_only %>% 
  mutate(disease = as_factor(disease), 
         author_affiliation_type = as_factor(str_replace(author_affiliation_type, ' ', '+'))
         ) %>% 
  mutate(collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'purely_academic', 
                                         'mixed')
                                 )
         ) %>% 
  count(collab_type, sort = TRUE)

knitr::kable(fmd_studies_by_collab_type)
```
## FMD by objective type
```{r echo=FALSE}
fmd_studies_by_objective_and_outbreak_type <- data_fmd_only %>% 
  count(objectives, outbreak_type) %>% 
  pivot_wider(names_from = outbreak_type, values_from = n)

@@

## Vaccine impact categories (FMD)
```{r echo=FALSE}
vax_impact_categories_fmd <- data_fmd_only %>%
  filter(vax_modelled_with_non_vax == 'yes') %>% 
  count(is_vax_effective, 
        name = 'num_of_studies', 
        sort = T
        )

@@

```{r echo=FALSE}
#' Count the number of each author affiliation type per year (including FMD)
collap_types_by_year <- data_without_fmd %>% 
  mutate(year_aggreg = as.numeric(ifelse(publication_year < 2006, 
                                         2005, 
                                         publication_year
                                         )
                                  ), 
         collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'Purely academic', 
                                         'Mixed')
                                 )
         ) %>% 
  count(year_aggreg, 
        collab_type, 
        name = 'total_publications'
        ) %>% 
  pivot_wider(names_from = collab_type, values_from = total_publications)

@@

```{r echo=FALSE}
academic_collabs_versus_others <- data_without_fmd %>% 
  filter(author_affiliation_type != 'academic_institutions') %>% 
  mutate(year_aggreg = as.numeric(ifelse(publication_year < 2006, 2005, publication_year
                                         )
                                  ), 
         collab_type = as_factor(if_else(str_detect(author_affiliation_type, 'academic_institutions'), 
                                         'Academic', 
                                         'Other'
                                         )
                                 )
         ) %>% 
  count(year_aggreg, 
        collab_type, 
        name = 'total_publications'
        )

@@

# Connection of authors to the studied locations 
```{r}
author_location_count_by_disease <- data_without_fmd %>%
  separate_rows(disease, sep = ' ') %>% 
  mutate(author_in_country_studied = as_factor(author_in_country_studied), 
         disease = as_factor(disease)
         ) %>% 
  group_by(collab_type, disease) %>%
  distinct(title, .keep_all = TRUE) %>% 
  count(author_in_country_studied, sort = TRUE) %>% 
  complete(author_in_country_studied, fill = list(n = 0)) %>% 
  pivot_wider(names_from = author_in_country_studied, values_from = n) %>% 
  replace_na(list(yes = 0, no = 0, not_applicable = 0)) %>% 
  mutate(total = not_applicable + no + yes) %>% 
  ungroup(collab_type) %>% 
  arrange(desc(total))
